Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change

### 2020-10-05 Scripts/Manuscript.Rmd
changes for thesis

# Abstract

@@

# Introduction

@@

# Methods

## Databases and species selection

@@

## Projecting species distributions under climate change

@@

## Changes in Stocks Share Ratio

@@

## Estimating the Time of Emergence

@@

## Vulnerability index

@@

## Caveats and uncertainties

@@

## Statistical analysis

@@

# Results

@@

## Changes in Stock Share Ratio (SSR)

@@

We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario. Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig. 1A). Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**). However, this result masks larger changes in SSR for specific species. For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%. By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR >< 1 s.d. of historical distribution will be relatively large (global mean of 69% of species in an EEZ). For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Fig. 1A-Polygons). Regions where EEZs are relatively small and have multiple borders (e.g. The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Fig. 1B). Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to the early $21^{st}$ century but are expected to increase in intensity (**Supplements Fig 1 (histogram)**).

```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005. Arrows go from the losing EEZ (point) to the winner (arrowhead). Pannel B zooms in to specific areas"}

@@

![Figure 1. Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead). Polygons represent the percentage of initial stocks that will see a shift. Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png)

## Time of Emergence

We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005). Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g. *Engraulis ringens* shared by Chile and Peru emerged in 2010). Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Fig. 2). The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Fig. 2).

```{r freq_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Cumulative increase in time of emergence by EEZ and stock"}

![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png)

Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Fig. 3). In a few EEZs stocks will have a relative late average ToE (e.g. Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their stocks emerged (e.g. average ToE across all stocks <= 2020). A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g. The Caribbean and the Mediterranean) (Fig 2). In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Fig. 2 - Polygons). However, in some cases emerging stocks represent 85% of total fishing revenue (e.g. Peru) or more (e.g. Marshall Islands, 92%). Only in very few cases like Brazil and Alaska's polar EEZ, no stock will emerge between 2006 and 2100.

```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}

@@

![Figure 3. Average time of emergence of transboundary stock. Land polygon represents  the current participation of emerging stocks in the country's fishing revenue. Exclusive Economic Zone polygon represents the average time of emergence across all stocks. Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png)

Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Fig. 4). This includes the United States who occupies the $4^{th}$ place. These 47 countries are also among those with the largest number of emerging taxa. Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Fig. 4). From the top ten countries with more revenue from transboundary species, nine of them share a border. Most of these countries (e.g.China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.

```{r bubble_plot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}

@@

![Figure 4. Average time of emergence (ToE) per fishing entity. Fishing revenue in logarithmic scale. Showing names for the top 75th percentile. Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png)

We compared species ToE within EEZs by geographic region and by species habitat preference (Fig. 5). The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001). Post-hoc tests revealed regional patterns. EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions. Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*). When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001). Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (*See Table Sup XX for full statistics*).

```{r Boxplot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}

@@

![Figure 5. Regional time of emergence per habitat preference. Regions acording to the United Nations and habitat preference to FishBase. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png)

We hypothesized that species environmental preferences (e.g. presence in certain biomes and habitats), climate change effects on water temperature (e.g. ToE of SST), human drivers (e.g. catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (*Table Sup 3*). Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table 1).

Table 1. Multiple regression results for time of emergence of transboundary stocks. Only showing significant variables. The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05).

```{r table1, eval = T,  echo = F}

read_csv("../Tables/mlr_result.csv") %>% 
  knitr::kable()

@@

# Discussion

In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species. Our findings here are noteworthy in several ways. First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Fig. 1). Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Fig. 2). Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Fig. 3).

By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig. 1). Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific. However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors. This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2]. Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig. 1B). These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in]. However, game theory suggest that the more number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.

We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs. Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig. 2). This is in alignment with previous research showing changes in the composition of marine catches since 1970's as evidence of climate change [@Cheung:2013gk]. Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk]. For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @@Zeidberg:2007gs]. In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij]. The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g. temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy]. It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 22.6% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled [@Rodggers:2015j].

Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries. For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move. The tropics are among the most species-rich biome [@Reygondeau:2019no] and have potentially experienced some of the largest declines in $O_2$ and greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id]. Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk]. The premise of a shift already happening is supported by observational evidence of tropical species apearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk]. In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relative small EEZs and many borders that match peaks of emerging stocks. Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Fig. 3).

An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Fig. 3). For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Fig. 5). Countries in these regions are highly dependent of pelagic-oceanic (e.g. tunas) and reef-associated transboundary species (e.g. *Speceies*) for both food security [@REF] and profits [@REF]. For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Fig. 5). In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that will impact important fisheries. However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability. While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portraited by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].

# Conclusion

@@

# Supplements

```{r fig_s1, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)"}

@@

![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png)
```{r fig_s2, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1970-2005)"}

@@

![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png)

```{r fig_s3, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}

@@

![Figure S3. Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead). Polygons represent the percentage of initial stocks that will see a shift. Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Supp3.png)

# Recycled  text
<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies. -->

>>>>

<!-- # Abstract -->

@@

## Introduction

@@

## Materials and methods

### Databases and species selection

@@

### Projecting species distributions under climate change

@@

### Changes in Stocks Share Ratio

@@

### Estimating the Time of Emergence

@@

### Vulnerability index

@@

### Caveats and uncertainties

@@

### Statistical analysis

@@

## Results

@@

### Changes in Stock Share Ratio (SSR)

@@

We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario. Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Figure \ref{fig:Figure1}A). Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**). However, this result masks larger changes in SSR for specific species. For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%. By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR $\pm$ 1 s.d. of historical distribution will be relatively large (global mean of 69% of species in an EEZ). For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Figure \ref{fig:Figure1}A-Polygons). Regions where EEZs are relatively small and have multiple borders (e.g. The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \ref{fig:Figure1}B). Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (Figure \ref{fig:FigureS1}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \ref{fig:FigureS2}).

```{r Figure1, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead). Polygons represent the percentage of initial stocks that will see a shift. Panel B zooms in to specific areas shown in grey boxes of A."}

img <- readPNG("../Figures/Fig1.png")
grid.raster(img)

<!-- ![Figure 1. Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead). Polygons represent the percentage of initial stocks that will see a shift. Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->

### Time of Emergence

We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005). Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g. *Engraulis ringens* shared by Chile and Peru emerged in 2010). Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Figure \ref{fig:Figure2}). The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Figure \ref{fig:Figure2}).

```{r Figure2, eval = T,  echo = F, fig.width = 5, fig.height=5, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock."}

@@

<!-- ![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png) -->

Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Figure \ref{fig:Figure3}). In a few EEZs stocks will have a relative late average ToE (e.g. Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their stocks emerged (e.g. average ToE across all stocks <= 2020). A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g. The Caribbean and the Mediterranean) (Figure \ref{fig:Figure2}). In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Figure \ref{fig:Figure2} - Polygons). However, in some cases emerging stocks represent 85% of total fishing revenue (e.g. Peru) or more (e.g. Marshall Islands, 92%). Only in very few cases like Brazil and Alaska's polar EEZ, no stock will emerge between 2006 and 2100.

```{r Figure3, eval = T,  echo = F, fig.width  = 12, fig.height =10, fig.cap =  "Average time of emergence of transboundary stock. Land polygon represents  the current participation of emerging stocks in the country's fishing revenue. Exclusive Economic Zone polygon represents the average time of emergence across all stocks. Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean)."}

@@

<!-- ![Figure 3. Average time of emergence of transboundary stock. Land polygon represents  the current participation of emerging stocks in the country's fishing revenue. Exclusive Economic Zone polygon represents the average time of emergence across all stocks. Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->

Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Figure \ref{fig:Figure4}). This includes the United States who occupies the $4^{th}$ place. These 47 countries are also among those with the largest number of emerging taxa. Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Figure \ref{fig:Figure4}). From the top ten countries with more revenue from transboundary species, nine of them share a border. Most of these countries (e.g.China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.

```{r Figure4, eval = T,  echo = F, fig.width=5, fig.height=6, fig.cap =  "Average time of emergence (ToE) per fishing entity. Fishing revenue in logarithmic scale. Showing names for the top 75th percentile. Countries marked with an asterisk have at least one neighbouring border."}

@@

<!-- ![Figure 4. Average time of emergence (ToE) per fishing entity. Fishing revenue in logarithmic scale. Showing names for the top 75th percentile. Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png) -->

We compared species ToE within EEZs by geographic region and by species habitat preference (Figure \ref{fig:Figure5}). The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001). Post-hoc tests revealed regional patterns. EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions. Other regions also significantly differ from each other but in a less frequent manner (See Table \ref{tab:TabS1}) for full statistics). When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001). Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TabS2} for full statistics).

```{r Figure5, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  " Regional time of emergence per habitat preference. Regions acording to the United Nations and habitat preference to FishBase. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers (outside of 1.5*IQR."}

@@

<!-- ![Figure 5. Regional time of emergence per habitat preference. Regions acording to the United Nations and habitat preference to FishBase. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->

We hypothesized that species environmental preferences (e.g. presence in certain biomes and habitats), climate change effects on water temperature (e.g. ToE of SST), human drivers (e.g. catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (Table \ref{tab:TableS3}). Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).

```{r Table1, eval = T,  echo = F, message = F, warning  = F}

read_csv("../Tables/mlr_result.csv") %>% 
  knitr::kable(
    "latex",
    booktabs = T,
    captio= "Multiple regression results for time of emergence of transboundary stocks. Only showing significant variables. The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
  )

@@

## Discussion

In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species. Our findings here are noteworthy in several ways. First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Figure \ref{tab:Figure1}). Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Figure \ref{tab:Figure2}). Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3}).

By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \ref{tab:Figure1}). Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific. However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors. This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2]. Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig. 1B). These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in]. However, game theory suggest that the more number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.

We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs. Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Figure \ref{tab:Figure2}). This is in alignment with previous research showing changes in the composition of marine catches since 1970's as evidence of climate change [@Cheung:2013gk]. Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk]. For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs]. In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij]. The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g. temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy]. It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 22.6% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled [@Rodggers:2015j].

Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries. For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move. The tropics are among the most species-rich biome [@Reygondeau:2019no] and have potentially experienced some of the largest declines in $O_2$ and greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id]. Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk]. The premise of a shift already happening is supported by observational evidence of tropical species apearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk]. In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relative small EEZs and many borders that match peaks of emerging stocks. Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Figure \ref{tab:Figure3}).

An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3}).. For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Figure \ref{tab:Figure5}). Countries in these regions are highly dependent of pelagic-oceanic (e.g. tunas) and reef-associated transboundary species (e.g. *Speceies*) for both food security [@REF] and profits [@REF]. For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Figure \ref{tab:Figure5}). In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that will impact important fisheries. However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability. While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portraited by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].

## Conclusion

@@

## Supplements

```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=6, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}

@@

<!-- ![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png) -->

```{r Figure2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}

@@

<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png) -->

```{r FigureS3, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead). Polygons represent the percentage of initial stocks that will see a shift. Panel B zooms in to specific areas shown in grey boxes of A."}

img <- readPNG("../Figures/Supp3.png")
grid.raster(img)

@@

```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98 confidenc teshold (yellow"}

img <- readPNG("../Figures/Supp4.png")
grid.raster(img)

<!-- # Recycled  text -->
<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies. -->
